---
title: "STA 445 HW2"
author: "Mason Nabbefeld"
date: "`r Sys.Date()`"
output: pdf_document
---
```{r setup, include = FALSE}
library(tidyverse)
```

## Problem 1

Create a vector of three elements (2,4,6) and name that vector `vec_a`. Create a second vector, `vec_b`, that contains (8,10,12). Add these two vectors together and name the result `vec_c`.
```{r}
vec_a <- c(2,4,6)
vec_b <- c(8,10,12)
vec_c <- c(vec_a + vec_b)
vec_c
```

## Problem 2

Create a vector, named `vec_d`, that contains only two elements (14,20). Add this vector to `vec_a`. What is the result and what do you think R did (look up the recycling rule using Google)? What is the warning message that R gives you?
```{r}
vec_d <- c(14,20)
c(vec_a + vec_d)
```

The result is a vector 16 24 20. I think that R did 14+2 for the first one and 4+20 for the second one, which is what its supposed to do, but then I think it went back to the first number in the smaller vector and added it to the last element in the larger vector.

## Problem 3

Next add 5 to the vector vec_a. What is the result and what did R do? Why doesn't in give you a warning message similar to what you saw in the previous problem?
```{r}
vec_a + 5
```

R did the correct thing in adding 5 to all elements in the vector, I think it didn't give me a warning because I was adding a constant to each element not adding elements together.

## Problem 4

Generate the vector of integers $\left\{ 1,2,\dots5\right\}$ in two different ways. 

a.  First using the `seq()` function
```{r}
seq(1,5)
```

b. Using the `a:b` shortcut.
```{r}
1:5
```

## Problem 5

Generate the vector of even numbers $\left\{ 2,4,6,\dots,20\right\}$ 

a. Using the seq() function  
```{r}
seq(2,20, by = 2)
```

b. Using the a:b shortcut and some subsequent algebra. 
```{r}
x <- 1:10
x*2
```

## Problem 6

Generate a vector of 21 elements that are evenly placed between 0 and 1 using the `seq()` command and name this vector `x`. 
```{r}
seq(0,1, length.out = 21)
```


## Problem 7

Generate the vector $\left\{ 2,4,8,2,4,8,2,4,8\right\}$ using the `rep()` command to replicate the vector c(2,4,8). 
  
```{r}
rep(c(2,4,8), 3)
```


## Problem 8
Generate the vector $\left\{ 2,2,2,2,4,4,4,4,8,8,8,8\right\}$
using the `rep()` command. You might need to check the help file for rep() to see all of the options that rep() will accept. In particular, look at the optional argument `each=`.

```{r}
rep(c(2,4,8), each = 3)
```

## Problem 9 

In this problem, we will work with the matrix 
$$ \left[\begin{array}{ccccc}
    2 & 4 & 6 & 8 & 10\\
    12 & 14 & 16 & 18 & 20\\
    22 & 24 & 26 & 28 & 30
    \end{array}\right]$$
    
a.  Create the matrix in two ways and save the resulting matrix as `M`.

i. Create the matrix using some combination of the `seq()` and `matrix()` commands.
```{r}
seq1 <- seq(2,30, by = 2)
M <- matrix(seq1,nrow = 3, byrow = TRUE)
M
```

ii. Create the same matrix by some combination of multiple `seq()` commands and either the `rbind()` or `cbind()` command.

```{r}
seq1 <- seq(2,10, by = 2)
seq2 <- seq(12,20, by = 2)
seq3 <- seq(22,30, by = 2)
M <- rbind(seq1,seq2,seq3)
M
```

b. Extract the second row out of M.
```{r}
M[c(2,5,8,11,14)]
```

c. Extract the element in the third row and second column of M
```{r}
M[6]
```


## Problem 10 

The following code creates a `data.frame` and then has two different methods for removing the rows with `NA` values in the column `Grade`. Explain the difference between the two.

```{r, eval=FALSE}
df <- data.frame(name= c('Alice','Bob','Charlie','Daniel'),
                     Grade = c(6,8,NA,9))
    
df[ -which(  is.na(df$Grade) ), ]

df[  which( !is.na(df$Grade) ), ]
```

The first one identifies the rows with NA values then excludes those rows from the data frame, and the second one identifies the rows without NA values and selects those.

## Problem 11 

Create and manipulate a list.

a.  Create a list named my.test with elements
        + x = c(4,5,6,7,8,9,10)
        + y = c(34,35,41,40,45,47,51)
        + slope = 2.82
        + p.value = 0.000131
```{r}
my.test <- list(x = c(4,5,6,7,8,9,10),
                y = c(34,35,41,40,45,47,51),
                slope = 2.82,
                p.value = 0.000131)
str(my.test)
```

b.  Extract the second element in the list.
```{r}
my.test[2]
```

c.  Extract the element named `p.value` from the list.
```{r}
my.test['p.value']
```



